---
title: 'Final Project: R Reference'
author: "Andrew Gordon"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

### Load a package that's already in R:
* library(the package you want)
* require(the package you want)
```{r}
## library()
library(palmerpenguins)

## require()
require(palmerpenguins)
```

### Load a data set into a dataframe:

1. require(here)
2. read.csv(here("folder where the data is","datafile.csv"))

```{r}
require(here)
gingko <- read.csv(here("data","ginkgo_data_2021.csv"))
```

## Data Structures

### c()

The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
* I can’t combine character and numeric types in the same call to c()
Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

```

Show the contents of the vector by typing its name in the console or using print()

```{r}
## Type the name of the vector to show the contents
num_vec
## [1]  1  4  8  9 13
## The print() function to show the contents:
print(char_vec)
## [1] "a"            "fish"         "data is cool"
```

### length()

Returns the number of elements within an R object

```{r}
## define a vector, list, or string
num_vec <- c(1, 5, 300, 64, 8, 232)
length(num_vec)
```

### matrix()
A matrix is an arrangement of data in a 2D structure

* must have the same data type

basic syntax:
  matrix(data, nrow, ncol, byrow)
  
  * data = vector of data for the matrix
  * nrow = number of rows
  * ncol = number of columns
  * byrow = set "= TRUE" if you want the data arranged by row

```{r}
num_matrix <- matrix(3:11,
                     nrow = 3,
                     ncol = 3,
                     byrow = TRUE)

print(num_matrix)
```

### data.frame()
a data frame can hold multiple outputs of data

```{r}
kids_educ.df <- data.frame(
  'alphabet' = c("a", "b", "c"),
  'numbers' = 1:3
)

print(kids_educ.df)
```
* use nrow() to count the number of rows in a data frame
* use ncol() to count the number of columns in a data frame
* use dim() to retrieve the number rows AND columns in a data frame (rows, columns)
```{r}
require(here)
gingko <- read.csv(here("data","ginkgo_data_2021.csv"))

nrow(gingko)
ncol(gingko)
dim(gingko)
```
## Subsetting
### $
Used to select  one of the columns within a data frame
```{r}
require(here)
gingko <- read.csv(here("data","ginkgo_data_2021.csv"))

g_depth <- gingko$max_depth
```

### [ ]
Used to select rows and columns in a data frame by ordered number
* syntax = data.frame[row #, column #]

```{r}
require(here)
gingko <- read.csv(here("data","ginkgo_data_2021.csv"))

#select first row
gingko[1,]

#select element in row 2, column 3
gingko[2,3]

#select third column
gingko[,3]
```

### subset()
You can use subset() to select a group of rows for a specific column value
* syntax: subset(data, column_name = "value")
```{r}
 require(palmerpenguins)
Adelie_penguins <- subset(penguins, species == "Adelie")
print(Adelie_penguins)
```

## Numerical Data Exploration

### summary() 
Used to obtain a a brief synopsis of the data 
```{r}
summary(penguins)
```

### mean()
Used to get the average for a collection of values
```{r}
mean(penguins$body_mass_g, na.rm = TRUE)
```

### sd()
Used to get the standard deviation for a variable in the dataset
```{r}
sd(penguins$body_mass_g, na.rm = TRUE)
```
## Graphical Data Exploration

### Scatterplot: plot()
arguments:

* col = color of the observations
* pch = character symbols for observations
* cex = scale of plotting characters
* main = title
* xlab = x axis label
* ylab = y axis label
* xlim = x limits of the plot
* ylim = y limits of the plot

```{r}
plot(gingko$max_depth, gingko$max_width,
     main = "Plot of Max Depth by Max Width",
     xlab = "Max Depth mm",
     ylab = "Max Width mm",
     col = "darkgreen",
     pch = 2,
     cex = 1,
     xlim = c(0,150),
     ylim = c(0,140)
     )
     
```

### Histogram: hist()
Same arguments as scatterplot PLUS:

* breaks()
* give the number of bars you want OR a vector of break points

```{r}
hist(penguins$flipper_length_mm, breaks = 12)
```

### Boxplot: boxplot()
when making conditional boxplots, use y ~ x as "y by x"
```{r}
require(here)
gingko <- read.csv(here("data","ginkgo_data_2021.csv"))

boxplot(gingko$petiole_length,
        main = "Boxplot of Petiole Length")
boxplot(max_width ~ seeds_present, data = gingko,
        main = "Boxplot of Petiole Length \n by Seeds Present")
```

### How to create four panel plots 
Put par(mfrow = ) BEFORE your plotting code
* Set mfrow = c(# of rows, # of columns)

```{r}
par(mfrow = c(2,2))
hist(gingko$petiole_length,
     main = "Histogram of Petiole Length",
     xlab = "Petiole Length mm",
     ylab = "Frequency")
hist(gingko$max_width,
     main = "Histogram of Max Width",
     xlab = "Max Width",
     ylab = "Max Width mm")
hist(gingko$max_depth,
     main = "Histogram of Max Depth",
     xlab = "Max Depth mm",
     ylab = "Frequency")
hist(gingko$notch_depth,
     main = "Histogram of Notch Depth",
     xlab = "Notch Depth mm",
     ylab = "Frequency")
```

## Distribution Functions

The first 3 functions are associated with $normal distributions. These functions use continuous distributions, so the term is "density"

The second 3 functions are associated with $binomial distributions. These functions use discrete distribution, so the term is "mass"

### dnorm()
Used to obtain the value of a Probability Density Function at a certain number

* aka, how often did x occur in this distribution?

arguments:

* x = entered number value
* mean = default of 0
* sd = default of 1

```{r}
dnorm(0)
```
### pnorm()
Cumulative Density Function

* aka, probability observing a value at or below x. To observe above x, use 1 - pnorm()
```{r}
pnorm(5, 5, 3)
```
### qnorm()
Quantile Function

* Gives the x value that indicates a certain percentile

```{r}
qnorm(0.5, 10, 5)

qnorm(c(0.025, 0.975))
```

### dbinom()
Probability Density Function

* used to find probability of observing x number of successes

arguments:
* x = number of successes
* size = number of trials
* prob = probability of success for each trial
```{r}
dbinom(
  x = 3,
  size = 6,
  prob = .4)


```
### pbinom()
Cumulative Density Function

* used to observe the probability of observing a value of x or less/or more

arguments:
* q = random variable
* size = number of trials
* prob = probability of success for each trial
* lower.tail = TRUE or FALSe. If you're looking to the right of q, use FALSE. to the left, use TRUE.

```{r}
pbinom(q = 3,
       size = 10,
       prob = 0.5,
       lower.tail = FALSE)
```

### qbinom()
Quantile Density Function

* Used to observe quantiles (i.e. 50th quantile, 90th quantile) where q is the quantile

```{r}
qbinom(.50,
       size = 12,
       prob = .6)

#The 50th quantile occurs at 7
```